A Framework for Discrete Change

### 2008-11-03 papers/invidividual/individual.tex
added papers directory which is not complete
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@232 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Reponse distributions and parameters}

% this needs some more on the design, interface to glm and similar
% other models etc. 
The package is built using S4 classes (object oriented classes in R)
to allow easy extensibility (REFERENCE to S4 methodology).

@@

Still to add models and data.

@@

@@

@@

@@

Ingmar Visser was supported by an EC Framework 6 grant, project 516542
(NEST).  Maarten Speekenbrink was supported by the ESRC Centre for
Economic Learning and Social Evolution (ELSE).  Thanks to Hilde
Huizenga, Brenda Jansen and Anna van Duijvenvoorde for the Iowa
Gambling task data.

>>>>

\subsection{Response distributions and parameters}

% this needs some more on the design, interface to glm and similar
% other models etc. 
The package is built using S4 classes (object oriented classes in R)
to allow easy extensibility (Chambers, 1998).

@@

%Still to add models and data.

The Weather Prediction Task (WPT, Knowlton, Squire \& Gluck, 1994) 
is a probabilistic categorization task, in which participants learn to 
predict the state of the weather (sunny, or rainy) on the basis 
of four ``tarot'' cards (cards with abstract geometrical patterns). 
Each cue pattern is associated with a particular probability distribution 
over the states of the weather. In order to perform in the task, 
participants must predict the weather in accordance with these
conditional probabilities.

%The WPT has been popular in neuropsychological research, 
%particularly because amnesic patients perform this task rather
%well, despite not being able to remember actually many aspects
%of the task (or in some cases, even performing the task at all).
%This has led to the conclusion that probabilistic category learning
%depends on implicit memory, which is separate from explicit
%memory. While this conclusion is debatable (Speekenbrink, Channon 
%\& Shanks, 2008), the finding of relatively unimpaired performance
%by amnesic individuals remains striking.

There are different accounts of probabilistic category learning. 
According to instance or exemplar learning theories, participants
learn by storing each encountered cue-outcome pairing. When presented
with a cue pattern, these exemplars are retrieved from memory, and weighted
according to their similarity to the probe cue pattern, to form a 
classification. According to associative theories, participants gradually 
learn by gradually associating the individual cues (or cue patterns in 
configural learning) to the outcomes. In rule-learning, participants are 
taken to extract rules by which to categorize the different cue patterns. 
Gluck, Shohamy and Myers (2002) proposed a number of such rules (or strategies). A main
difference between these is whether responses are based on the 
presence/absence of a single cue, or whether responses are based on 
cue patterns. Gluck et al. formulated all strategies in a deterministic and
optimal manner (e.g., the multi-cue strategy corresponded to giving the optimal
response to each cue pattern). Meeter et al. allowed for probabilistic 
responding (a small probability of giving the non-optimal response).

Alternative non-strategy based analyses of the WPT (Lagnado et al, Speekenbrink et al) 
have estimated response strategies by variations of logistic regression.

%associative, rule-based.  
Here, we analyze the behavior of a single individual performing the
WPT for 200 trials. We let each state be characterized by a Generalized
Linear Model with a Bernoulli response and logistic link function. We are
interested in whether a DMM can recover a strategy model in line with Gluck et al. 
As we fit the data to a single subject, we must place some constraints.
Specifically, we constrain the state transitions to be in a ``left-right'' format
(states can only proceed to the immediately adjacent state and never back) and 
the initial state

A single state model (usual GLM) 
We started with a 3-state model,

@@

Chambers, J. M. (1998). Programming with Data: A Guide to the S Lan-
guage. New York: Springer-Verlag.

@@

Gluck, M. A., Shohamy, D., \& Myers, C. (2002). How do people solve the 
weather prediction task?: Individual variability in strategies for 
probabilistic category learning. Learning \& Memory, 9, 408-418.

@@

Knowlton, B. J., Squire, L. R., \& Gluck, M. A. (1994). 
Probabilistic classification learning in amnesia. 
Learning \& Memory, 1 , 106-120.

@@

Ingmar Visser was supported by an EC Framework 6 grant, project 516542
(NEST).  Maarten Speekenbrink was supported by the ESRC Centre for
Economic Learning and Social Evolution (ELSE).  Thanks to Hilde
Huizenga, Brenda Jansen and Anna van Duijvenvoorde for the Iowa
Gambling task data and David Lagnado for the Weather Prediction Task
data.
